---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# excelcompare

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![R-CMD-check](https://github.com/SchmidtPaul/excelcompare/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/SchmidtPaul/excelcompare/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

Compare Excel files (.xlsx) cell by cell. Unlike standard table comparison tools, excelcompare captures **every cell** in a worksheet - including titles, footnotes, and annotations that live outside the main data table.

## Installation

You can install the development version of excelcompare from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("SchmidtPaul/excelcompare")
```

## Example

```{r example}
library(excelcompare)

# Paths to example Excel files shipped with the package
path1 <- system.file("extdata", "example1.xlsx", package = "excelcompare")
path2 <- system.file("extdata", "example2.xlsx", package = "excelcompare")

# Compare the two files
diff <- compare_xlsx(path1, path2)

# The result is a tibble with one row per difference
diff
```

```{r print}
# The print method shows waldo-inspired diffs
print(diff)
```

## Features

- **Cell-level comparison** - Detects differences in any cell, not just tabular data
- **Numeric tolerance** - `tolerance` parameter for floating-point comparison
- **NA-equals-zero** - `na_equals_zero` parameter to treat empty cells and zeros as equal
- **Multi-sheet support** - Compare specific or all shared sheets (`sheet` parameter)
- **Clean diff output** - S3 print/summary methods with waldo-inspired formatting
- **`safe_to_numeric()`** - Exported utility for robust numeric conversion (handles European formats, thousand separators)

## Learn more

- [Getting Started vignette](https://schmidtpaul.github.io/excelcompare/articles/excelcompare.html)
- [Function reference](https://schmidtpaul.github.io/excelcompare/reference/)

Built on [{tidyxl}](https://nacnudus.github.io/tidyxl/) for comprehensive Excel cell reading. Only supports modern Excel formats (.xlsx, .xlsm).
