---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# excelcompare

<!-- badges: start -->
[![R-CMD-check](https://github.com/SchmidtPaul/excelcompare/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/SchmidtPaul/excelcompare/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

Compare Excel files (.xlsx) cell by cell. Unlike standard table comparison tools, excelcompare captures **every cell** in a worksheet - including titles, footnotes, and annotations that live outside the main data table.

## Installation

You can install the development version of excelcompare from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("SchmidtPaul/excelcompare")
```

## Usage

```{r example, eval = FALSE}
library(excelcompare)

# Compare two Excel files
differences <- compare_xlsx("file1.xlsx", "file2.xlsx")

# Compare specific sheets
differences <- compare_xlsx("file1.xlsx", "file2.xlsx", sheet = "Summary")
```

## Features

- **Cell-level comparison**: Detects differences in any cell, not just tabular data
- **Position-aware**: Reports exact row and column of each difference
- **Sheet selection**: Compare specific worksheets by name or index
- **Clean output**: Returns a tidy tibble of all differences

## Technical Details

Built on [{tidyxl}](https://nacnudus.github.io/tidyxl/) for comprehensive Excel cell reading. Only supports modern Excel formats (.xlsx, .xlsm) - legacy .xls files are not supported.
