---
title: "Getting Started with excelcompare"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with excelcompare}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

**excelcompare** compares Excel files (.xlsx/.xlsm) at the cell level.
Unlike tools that only compare tabular data, excelcompare captures
*every cell* in a worksheet — including titles, footnotes, and
annotations outside the main data table.

```{r setup}
library(excelcompare)
```

## Basic Comparison

The main function is `compare_xlsx()`. It takes two file paths and
returns a tibble of cell-level differences.

```{r basic}
file1 <- system.file("extdata", "example1.xlsx", package = "excelcompare")
file2 <- system.file("extdata", "example2.xlsx", package = "excelcompare")

diffs <- compare_xlsx(file1, file2)
diffs
```

Each row represents one cell that differs between the two files:

- **sheet**: the worksheet name
- **address**: the Excel cell address (e.g., "A1", "B3")
- **value1** / **value2**: the cell content in file 1 and file 2
- **status**: `"modified"` (changed), `"added"` (only in file 2), or
  `"removed"` (only in file 1)

## Numeric Tolerance

By default, values are compared as exact strings. If you expect small
numeric differences (e.g., from rounding), use the `tolerance` parameter:

```{r tolerance}
# Treat differences up to 0.05 as equal
diffs_tol <- compare_xlsx(file1, file2, tolerance = 0.05)
diffs_tol
```

Tolerance-aware comparison automatically handles:

- Numeric cells stored as numbers
- Text cells that look like numbers (including comma-decimal like `"10,5"`)
- Scientific notation (e.g., `"1e-5"`)

## NA-equals-Zero

In some workflows, empty cells and zero-valued cells are treated as
equivalent. Enable this with `na_equals_zero = TRUE`:
```{r na-zero}
compare_xlsx(file1, file2, na_equals_zero = TRUE)
```

## Multi-Sheet Comparison

Use the `sheet` parameter to control which sheets are compared:

```{r multi-sheet}
multi1 <- system.file("extdata", "example_multi1.xlsx", package = "excelcompare")
multi2 <- system.file("extdata", "example_multi2.xlsx", package = "excelcompare")

# Compare all shared sheets (sheet = NULL)
all_diffs <- compare_xlsx(multi1, multi2, sheet = NULL)
all_diffs
```

You can also specify sheets by name or index:

```{r specific-sheets}
# By name
compare_xlsx(multi1, multi2, sheet = "Data")

# Multiple sheets
compare_xlsx(multi1, multi2, sheet = c("Data", "Summary"))
```

## Summary View

For a quick overview, use `summary()`:

```{r summary}
summary(all_diffs)
```

## Utility: safe_to_numeric()

The package also exports `safe_to_numeric()`, which converts character
values to numeric while handling multiple decimal separators:

```{r safe-numeric}
safe_to_numeric("10.5")
safe_to_numeric("10,5")
safe_to_numeric("1.234,56")
safe_to_numeric("1e-5")
safe_to_numeric(c("10.5", "10,5", "text"))
```
